Dark matter halo shape in the Auriga simulations: radial dependence, time evolution and baryon disk alignment

### 2019-07-17 Document/paper.tex
edit

\begin{figure*}
\begin{center}
\includegraphics[width=0.8\textwidth]{correlation_T_MHD_disk.pdf}
\end{center}
\caption{Correlations between the halo triaxility at different radii
  and baryonic disk properties. 
  The label with the $\rho$ value corresponds to the Spearman's rank
  correlation coefficient.
  The line is the best linear minimum squares fit.
  The x-axis in the first column is the gas density at the center of
  the galaxy with in a sphere of radius  $1$ kpc \citep{Pakmor17};
  the second column shows the disk to total mass ratio and the last
  column includes the disk optical radius defined to be the radius at which the
  $B$-band surface brightness drops below 25 mag arcsec$^-2$ \citep{auriga}.
  The largest correlations are found for the two smaller radii and
  dilute as one approached $R_{200}$.
  Large and massive stellar disks with a low gas content are
  correlated with low dark matter triaxilities.}
\label{fig:disk_correlations}
\end{figure*}

@@

\section{Discussion and Conclusions}

Following that linea reasoning, one would also expect that the rounding
effect of baryons is related to some  galactic parameters such as its
component masses and radii. 
We look for these kind of correlations and to find that the strongest
can be found for the stellar density.
However, the correlation is relatively low ({\bf cuanto vale el
  coeficiente de correlation}) most likely due to the different
formation histories.
In other words, the effect of the baryonic disk on the shape of the DM halo
does not fully explain the deviation from oblateness of MHD halos at
$r<10$kpc.

{\bf Finalmente cuales son las cantidades disponibles en el disco?
  cuales son las que muestran una mayor correlacion? Haria falta un
  plot con todas las correlaciones y la mencion de alguna prueba de
  machine learning del fit lineal para ver cuales son los parametros
  mas importantes.}

  We measured the sphericity of a halo shape defined as the distance 
  from the point that characterizes the shape in the triaxiality plane
  to the point $(q=1,s=1)$, which represents a perfect sphere. Taking this 
  as our metric of interest, the sphericity question becomes a search for
  important variables and tendencies that affect this quantity. For this case,
  given the small sample of 30 halos, we chose to study this problem with different 
  3 methods to account for the lacking statistical consistency. We analyzed the effect
  of the Star disk Radius, Baryonic fraction, Gas density, BH density, Gas Radius and
  Star density.\\

  We started with a simple linear Lasso regression to force a classification of 
  important variables. Then, we converted this problem in a binary classification problem
  by partitioning the samples in two balanced groups according to the studied metric which was then fitted using Random forest.
  Finally, we concluded this analysis with an MCMC sampling of the likelyhood of a linear model
  with varying slope and intercept parameters.\\

  The results of each method are presented below:

\textit{ Talk about source of triaxiality at the inner parts of the
  halos (bar?). This source of triaxiality at the inner parts explains
  why the axial ratios are $\approx 0.95$ and not exactly $1$. We
  should also discuss that the decrease in the axial ratios for bigger
  radii may actually be bigger/steeper but it is dimmed by the
  contribution of inner parts.}

%\begin{figure}
%  \includegraphics[width=\columnwidth]{./pics/Delta asphericity Vs Star Density.png}
%  \caption{Difference in asphericities between MHD and DM shapes Vs
%    Star Density of the simulation. Unsure about this graphic. Take
%    delta asphericity as the strength of the rounding effect of
%    baryons.}  
%  \label{fig:Star_Density_effect}
%\end{figure}

Half of the observational constraints are computed in terms of
isodensity contours, the other half in terms of isopotential contours.
To compare our results against the second kind we must either
translate the isodensity results or recompute in terms of isopotential
regions.
For this purpose, we run a simple iterative algorithm to find an
approximation of the shape of the isopotential contour.

First, we calculate the mean and standard deviation of the potential over a
spherical shell of width equals to $10\%$ of the radius at which it is
sampled. 
Then, we calculate the inertia tensor of particles with potential
within $1\sigma$ around the mean potential and calculate its triaxial
characterization with the reduced inertia tensor. 
We repeat the
process of calculating the potential mean and standard deviation until
convergence is achieved with tolerance of $10^{-4}$.

In Figure \ref{fig:density_potential} we compare the two approaches to
measure the axial rations and plot the analytic expectation for $q$, 
$(1-q_{\phi})\approx \frac{1}{3}(1-q_{\rho})$
\citep{Binney_and_Tremaine_2008}, taking the volume-enclosed axial
ratios  as an approximation for the isodensity contour ratios
$q_{\rho}$. 
This plot is computed at four different radii.

Although the analytic expression is meant to be used on the outer
regions of logarithmic axisymmetric halos, it works well as a first
approximation for nearly axisymmetric halos as those produced by our
simulations 
We find that the difference between the measured and the
approximated isopotential axial ratios is not bigger than $quantity
percent$.

@@

@@

>>>>

\begin{figure*}
\begin{center}
\includegraphics[width=0.8\textwidth]{correlation_T_MHD_disk.pdf}
\end{center}
\caption{Correlations between the halo triaxility at different radii
  and baryonic disk properties. 
  The label with the $\rho$ value corresponds to the Spearman's rank
  correlation coefficient.
  The line is the best linear minimum squares fit.
  The x-axis in the first column is the gas density at the center of
  the galaxy with in a sphere of radius  $1$ kpc \citep{Pakmor17};
  the second column shows the disk to total mass ratio and the last
  column includes the disk optical radius defined to be the radius at which the
  $B$-band surface brightness drops below 25 mag arcsec$^{-2}$ \citep{auriga}.
  The largest correlations are found for the two smaller radii and
  dilute as one approached $R_{200}$.
  Large and massive stellar disks with a low gas content are
  correlated with low dark matter triaxilities.}
\label{fig:disk_correlations}
\end{figure*}

@@

@@

The first effect that we put in evidence in this paper is the effect
of baryons in producing rounder DM halos.
This effect has been already reported in other simulations \citep{Debattista08, Bryan13,
  Butsky16, Chua19, Artale19}.
The strength of the change depends on the numerical resolution, the
method to resolve the hydrodynamics and the models describing star
formation and stellar feedback.
The key concept behind these results is that the baryon distribution
influences and correlates with the dark matter halo shape.

In order to explore this idea in the Auriga simulations we looked for
correlations with stellar disk properties. 
Looking into the measurements already reported by \cite{auriga} and
\cite{pakmor17} we find three baryonic quantities that have the
strongest correlation with DM halo triaxiality: the central gas
density in a sphere of radius $1$kpc, the disk to total mass ratio and
the optical radius.

Figure \ref{fig:disk_correlations} summarizes the correlations of
these quantities with the triaxility at five diferent radii.
We ues the Spearmen's rank correlation coefficient to quantify the
correlation strength.

\citep{Debatistta08} shape aligment with respects to the disk changes
but saturates beyond the disk radius.

@@

\section{Conclusions}
\section{Discussion and Conclusions}
